The forwarding unit is responsible for forwarding newly computed data from later stages in the pipeline to earlier stages, where new instructions might need the data before they are committed to registers at the end of the pipeline.
The forwarding unit is essential to achieving a high performance pipelined design, as without it computations would either be wrong or slow.

The forwarding unit looks at which data is needed in the execution stage, and if the needed data has been updated down the pipeline but not in the register file, forwards the data to the ALU instead of the old data fetched from the register file by the instruction decode stage.
To be able to do this, the forwarding unit needs data from many different stages at once.

The forwarding unit uses the Hazard Detector component for the actual hazard detection, and performs its forwarding based on the output of the Hazard Detector.

\subsubsection{In Signals}

\begin{description}
\item{\textbf{Rs Address from ID/EX }} \\
    The source register address of the Rs operand from the ID/EX pipeline registers.

\item{\textbf{Rt Address from ID/EX}} \\
    The source register address of the Rs operand from the ID/EX pipeline registers.

\item{\textbf{Register Destination from EX/MEM}} \\
    The destination register address from the EX/MEM pipeline registers.

\item{\textbf{Register Destination from MEM/WB}} \\
    The destination register address from the MEM/WB pipeline registers.

\item{\textbf{Register Write from EX/MEM}} \\
    The register write control signal from the EX/MEM pipeline registers.

\item{\textbf{Register Write from MEM/WB}} \\
    The register write control signal from the MEM/WB pipeline registers.
\end{description}

\subsubsection{Out Signals}

\begin{description}
\item{\textbf{Forward Rs}} \\
    The forwarded value of the Rs register.

\item{\textbf{Forward Rt}} \\
    The forwarded value of the Rt register.

\end{description}
