This section describes the methodology used in creating the solution.

\section{Project Setup}

The Xilinx ISE project was set up with the design properties as in table \vref{table:design-properties}.

\begin{table}
    \begin{center}
        \begin{tabular}{ | l | l | }
            \multicolumn{1}{ l }{\textbf{Property Name}} &
            \multicolumn{1}{ l }{\textbf{Value}} \\
            \multicolumn{2}{ l }{ } \\
            \hline
            Top-Level Source Type & HDL \\
            \hline
            Family & Spartan6 \\
            Device & XC6SLX16 \\
            Package & CSG324 \\
            Speed & -3 \\
            \hline
            Synthesis tool & XST (VHDL/Verilog) \\
            Simulator & ISim (VHDL/Verilog) \\
            Preferred Language & VHDL \\
            Property Specification in Project File & Store non-default values only \\
            Manual Compile Orer & (not checked) \\
            VHDL Source Analysis Standard & VHDL-93 \\
            \hline
            Enable Message Filtering & (unchecked) \\
            \hline
        \end{tabular}
        \caption{Xilinx ISE Project Design Properties}
        \label{table:design-properties}
    \end{center}
\end{table}

\section{Tools}

This section contains a description list of the tools used for the assignment.

\subsection{Software}
\begin{description}
    \item{\textbf{ISE Project Navigator 12.4 (nt64) M.81d, expired licence}} \\
        Main IDE for writing VHDL.
    \item{\textbf{ISim 12.4 (nt64) M.81d, expired license}} \\
        Main simulation environment for simulating VHDL.
    \item{\textbf{ModelSim SE 6.6d}} \\
        Secondary simulation environment for simulating VHDL.
    \item{\textbf{Xilinx Platform Studio 12.4 (nt64) Build EDK\_MS4.81d+1, expired licence}} \\
        Used for preparing compiled VHDL for the FPGA board.
    \item{\textbf{Avnet Programming Utility}} \\
        Used for configuring the FPGA.
    \item{\textbf{Text editors}} \\
        Sublime Text 2, Vim 7.3, Notepad (©Copyright Microsoft Corporation).
    \item{\textbf{GNU command-line tools}} \\
        Grep, sed, find, etc.
    \item{\textbf{git 1.8.1.2}} \\
        Version control system.
    \item{\textbf{GitHub}} \\
        Remote code repository hosting, issue tracking, wiki for logging.
\end{description}

\subsection{Hardware}
\begin{description}
\item{\textbf{Xilinx Spartan-6 XC6SLX16 CSG324}}
    FPGA board.
\item{\textbf{Development PC, Windows 7}}
    For development.
\item{\textbf{USB cable}}
    For connecting the FPGA board to the development computer.
\end{description}

\bigskip
